Convergence rate of MAP estimates \\ for the exponential family

### 2021-04-14 main.tex
Upper bound for variance and mixed terms of the example

@@

@@

@@

\begin{example}
	For the gaussian variance MAP, the co-variance term is
	\begin{align}
		\frac{n}{n+n_0} \Cov(\hat \natp_n ; T(X) )  
		&= -\frac{n}{2} \expect{ \frac{X_0^2 - \mu^*}{\sum_i T(X_i)  + n_0 \mu_0} } \\
		&=  \half \expect{ \frac{n - \chi^2(n)}{\chi^2(n) + n_0 \frac{\mu_0}{\mu^*}}} \\
		& = \half[n] \expect{\inv{\chi^2(n) + a}} - \half \expect{\inv{1 + \frac{a}{\chi^2(n)}}}
	\end{align}
	where $a=n_0 \frac{\mu_0}{\mu^*}$, and at the second line we re-introduced the $X_i$ to form a chi-square. 
	These expectations are not simple to solve, but using wolfram alpha, we get  some formulas :
	\begin{align}
		\expect{\inv{\chi^2(n) + a}}  = \half e^{\frac{a}{2}} \text{ExpIntegralE}[\half[n],\half[a]]
	\end{align}
	where the \href{https://reference.wolfram.com/language/ref/ExpIntegralE.html}{exponential integral function} is defined as 
	\begin{align}
		\text{ExpIntegralE}[n , z] = E_n(z) = \int_1^\infty \frac{e^{-z t} }{t^n} dt = z^{n-1} \Gamma(1-n , z)
	\end{align}
	with the upper incomplete gamma function.
	Now our goal is to find a simple upper bound on this function.
\end{example}

@@

@@

>>>>

\newcommand{\MAPm}{\hat \mu_n}
\newcommand{\MAPt}{\hat \natp_n}

@@

\begin{example}
		We need a formula  for $\E[\MAPt]$ the expectation of the natural MAP estimate
	\begin{align}
		\E[\MAPt] &= -\frac{n+n_0}{2}\expect{\inv{n_0 \mu_0 + \sum_i X_i^2}}  \\
		&= -\frac{n+n_0}{2 \mu^*}\expect{\inv{n_0 \frac{\mu_0}{\mu^*}+ \sum_i \frac{X_i^2}{ \mu^*} }} \\
		&= -\frac{n+n_0}{2 \mu^*}\expect{\inv{a + \chi^2(n)}}
	\end{align}
	where $\sum_i \frac{X_i^2}{ \mu^*}$ is a chi-square random variable of degree $n$ and 
	\begin{align}
			a=n_0 \frac{\mu_0}{\mu^*} \; .
	\end{align}
	This expectations is not simple to solve, but using wolfram, we get  the formula :
	\begin{align}
		\expect{\inv{a + \chi^2(n)}}  = \half e^{\frac{a}{2}} E_{\half[n]}(\half[a])
	\end{align}
	where the \href{https://dlmf.nist.gov/8.19}{generalized exponential integral function} is defined as 
	\begin{align}
		E_k(z) = \int_1^\infty \frac{e^{-z t} }{t^k} dt \; .
	\end{align}
	Now our goal is to bound this function with simpler functions.
	Fortunately, mathematicians have been working on these integrals for decades.
	When $n=2$, we have the simple exponential integral function which \href{https://dlmf.nist.gov/6.8#E1}{verifies the bound}
	\begin{align}
		\half \log(1 + \frac{2}{x}) 
		\leq e^x E_1(x) 
		\leq \log(1+\inv{x}) \leq \inv{x} \; .
	\end{align}
	For larger $n$, we have \href{https://dlmf.nist.gov/8.19#E21}{the general bound}
	\begin{align}
		\inv{x+n}
		\leq e^x E_k(x) 
		\leq \inv{x + k - 1}
	\end{align}
	so that 
	\begin{align}
		\inv{a+n}
		\leq \expect{\inv{a + \chi^2(n)}}
		\leq \inv{a + n - 2}
	\end{align}
	so that
	\begin{align}
		\frac{n +n_0}{2 \mu^* (a + n)}	
		&\leq - \E[\MAPt]
		\leq \frac{n +n_0}{2 \mu^* (a + n - 2)}	\\
		\iff \inv{2 \mu_n}
		&\leq - \E[\MAPt]
		\leq \inv{2 \mu_n - \frac{4 \mu^*}{n + n_0}}
		\label{eq:theta_expectation_bound}
	\end{align}
	because $\frac{\mu^* (a + n)}{n +n_0} = \mu_n$. Remark how we almost bounded the deviation between $\E[\MAPt]$ and $\natp_n$.
\end{example}

	\begin{figure}[ht]
    	\centering
	    \includegraphics[width=.7\textwidth]{figs/wolfram.png}
    	\label{fig:wolfram}
	\end{figure}

@@

\begin{example}
	\begin{align}
		\bregmanconj(\mu^*; \mu_n)
		 &= \frac{\mu^*}{\mu_n} -1  - \log \frac{\mu^*}{\mu_n}  \\
		 & = \frac{n+n_0}{a + n} - 1 - \log \frac{n+n_0}{a +n} \\
		 & = \frac{n_0 - a}{a + n} - \log \frac{n+n_0}{a +n} \\		 
	\end{align}
\end{example}

@@

\begin{example}
	For the gaussian variance MAP, the variance term is
		\begin{align}
			\Cov(\hat \natp_n ; \hat \mu_n )  
			= \E[\MAPt \MAPm] - \E[\MAPt] \E[\MAPm] 
			=  - \half - \E[\MAPt] \mu_n
	\end{align}
	where we used the fact that $\natp \mu = -\half$ for all valid pairs $(\natp, \mu)$. 
	Plugging-in \eqref{eq:theta_expectation_bound} yield
	\begin{align}
		-\half + \frac{\mu_n}{2 \mu_n}
		&\leq \Cov(\hat \natp_n ; \hat \mu_n ) 
		\leq -\half + \half \frac{a + n}{a + n - 2} \\
		\iff 0&\leq \Cov(\hat \natp_n ; \hat \mu_n ) 
		\leq \half \frac{a+n - (a + n -2)}{a + n - 2} \\
		\iff 0&\leq \Cov(\hat \natp_n ; \hat \mu_n ) 
		\leq \inv{a +n - 2}
	\end{align}
	which is exactly the kind of upper bounds we were looking for !
\end{example}

@@

\begin{example}
	As we have the tight bounds \eqref{eq:theta_expectation_bound} for the expectation of the natural parameter of the MAP, we can bound above and below this mixed term
	\begin{align}
		0
		\leq \natp_n - \E[\MAPt]
		\leq \frac{n_0 +n}{2 \mu^*} \left ( \inv{a+n-2} - \inv{a+n} \right)
		 = \frac{n_0+n}{\mu^*(a+n)(a+n-2)}	 \in O(n^{-1}) \; .
	\end{align} 
	Remark how this asymptotic class differs from the $O(n^{-2})$ I expected. Is this bound less tight than I thought, or are my calculus wrong ? 
	Also remark that $\natp_n$ is always greater than $\E[\MAPt]$, which is a good sanity-check because $\nabla \conj(\mu) = - \inv{2\mu}$ is a concave function.

	The rate on the mixed term depends on the sign of $\mu^* - \mu_0 $
	\begin{align}
		\frac{n_0}{n+n_0} \langle  \mu^* - \mu_0 ; \natp_n - \E[\hat \natp_n] \rangle
		\leq \begin{cases}
			0 \ \text{ if } \ \mu^*\leq \mu_0, \\
			\frac{n_0 - a}{(a+n)(a+n-2)} \ \text{ otherwise.}
		\end{cases}
	\end{align}
\end{example}

@@

\citet{ostrovskii2021finite}  shows a convergence rate for the MLE (or other $M$-estimators)   when  the log-likelihood is self-concordant. This convergence rate holds only after a given number of samples have been seen. This is because they use the fact that self-concordant losses are upper bounded by a quadratic in a neighborhood of the optimum.
